An Object-Logic Explanation for the Inconsistency in G\"odel's Ontological Theory (Extended Abstract)

### 2016-07-11 Papers/2016/KI/main.tex

Scott's version of G\"odel's argument was automatically reconstructed by the higher-order theorem prover \textsc{Leo-II} \cite{C40} and its correctness was verified step-by-step in the \textsc{Coq} proof assistant \cite{CSR}. To bridge the gap between higher-order logics (HOL; cf.~\cite{andrewsSEP} and the references
therein), as used by these systems, and higher-order \emph{modal} logics (HOML; cf.~\cite{homl} and the
references therein), on which the ontological argument is based, the logic embedding approach \cite{J23,C40} was used.

However, G\"odel's original axioms, as used in his
manuscript \cite{GoedelNotes}, are inconsistent. This fact has
remained unnoticed to philosophers until 2013, 
when \textsc{Leo-II} found a surprising refutation of the axioms.
In \cite{C55} we extracted from Leo-II's machine-oriented refutation an informal and human-oriented intuitive explanation for the inconsistency, and we reconstructed and
verified it in the \textsc{Isabelle} proof assistant. But that explanation relied on reasoning at the \emph{meta-logic} (HOL) level, which was only possible because of the embedding. Here we complement that work with a purely \emph{object-logic} (HOML) explanation, and we compare and formalize both explanations in the \textsc{Coq} proof assistant.

>>>>

Scott's version of G\"odel's argument was automatically reconstructed
by higher-order automated theorem provers \cite{C40} and its
correctness was verified step-by-step in the \textsc{Coq} proof
assistant \cite{CSR}. To bridge the gap between higher-order logics
(HOL; cf.~\cite{andrewsSEP} and the references therein), as used by
these systems, and higher-order \emph{modal} logics (HOML;
cf.~\cite{homl} and the references therein), on which the ontological argument is based, the logic embedding approach \cite{J23,C40} was used.

However, G\"odel's original axioms, as used in his
manuscript \cite{GoedelNotes}, are inconsistent. This fact has
remained unnoticed to philosophers until 2013, 
when \textsc{Leo-II} \cite{J30} found a surprising refutation of the axioms.
In \cite{C55} we extracted from Leo-II's machine-oriented refutation an informal and human-oriented intuitive explanation for the inconsistency, and we reconstructed and
verified it in the \textsc{Isabelle} proof assistant. But that explanation relied on reasoning at the \emph{meta-logic} (HOL) level, which was only possible because of the embedding. Here we complement that work with a purely \emph{object-logic} (HOML) explanation, and we compare and formalize both explanations in the \textsc{Coq} proof assistant.
